---
title: "CobbDouglas"
author: "Jaime Duran"
date: "5/11/2020"
output: html_document
---

## non-linear optimization

This Rmarkdown file demonstrates techniques for non-linear optimization, and is illustrated with data for the Cobb-Douglas production function, $y = A K^\alpha L^\beta$, K capital, L labor, and the exponents the elasticies of production by input.

### Methods demonstrated:
 1. Grid search of space of parameters  
 2. non-linear optimization software, find values of the parameters that minimize the sum of squared errors, errors are (data-predicted vs actual data)
 3. non-linear optimization software, maximum likelihood- find values of the parameters that maximize the likelihood of the data assuming errors are normally distributed  
 4. log-transform to a linear model  
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in the data

```{r}
CobbDoug = read.csv('CobbDouglasData.csv')
attach(CobbDoug)
```


```{r}
plot(Cap, Q, pch=19, col='red', xlab= 'input values', ylab = 'output',
     main='Cobb-Douglas Production function')
points(Emp, Q, pch=19, col='blue')
grid()
```

A two-dimensional plot would be nice..
Set up a non-linear loss function

```{r}
CDloss=function(parms, Q, Cap, Emp){
  A = parms[1]
  alpha = parms[2]
  beta = parms[3]
  
  Loss = 0
  for (i in 1:length(Q)){
    Loss = Loss + ((Q[i]-A*(Cap[i]^beta*Emp[i]^alpha))^2)
  }
  
  return (log(Loss))
}

CDmle=function(parms, Q, Cap, Emp){
  A = parms[1]
  alpha = parms[2]
  beta = parms[3]
  
  logLoss = 0
  for (i in 1:length(Q)){
    errorTerm= (Q[i]-A*(Cap[i]^beta*Emp[i]^alpha))
    logLoss = logLoss - dnorm(errorTerm, log=T)
  }
  
  return (logLoss)
}
```

Test the loss function

```{r}
A = 1
alpha = 0.5
beta = 0.5
parms=c(A, alpha, beta)
print(CDloss(parms, Q, Cap, Emp))

# better or worse than
A = 0.5
alpha = 0.2
beta = 0.8
parms=c(A, alpha, beta)
print(CDloss(parms, Q, Cap, Emp))
```

## Grid search
  not the best idea, can get close to an answer 

```{r}

for (A in seq(from=0.01, to = 2, length.out = 4))
  for(a in seq(from=0.01, to = 1,length.out = 6))
   for(b in seq(from=0.01, to = 1, length.out = 6)){
     parms = c(A, a, b)
      print(paste(A, a, b,'        ', CDloss(parms, Q, Cap, Emp)))
  }

```

] "1.33666666666667 0.604 0.406          14.2090301971589"

## non linear optimization

use function optim to find values of A, alpha, beta
find values that minimize the loss function, loss = sum of sqrared errors.  This is a 'least squares' solution

```{r}
##     ?optim
parms = c(2, 0.8, 0.2)
opt = optim(parms, CDloss, Q=Q, Cap=Cap, Emp=Emp)
print(opt)
parms = c(opt$par[1], opt$par[2], opt$par[3])
print(CDloss(parms, Q, Cap,Emp))
```

## maximum likelihood estimation

With the right fit, the residual errors should be independent, normally distributed, with mean 0. likelihood of the parameter set (a, $\alpha$, $\beta$) is

$\prod_0^n \phi(e_i)$

$\phi$ is the standard normal distribution


```{r}
parms = c(2, 0.8, 0.2)
opt = optim(parms, CDmle, Q=Q, Cap=Cap, Emp=Emp)
print(opt)
parms = c(opt$par[1], opt$par[2], opt$par[3])
print(CDmle(parms, Q, Cap,Emp))
print(CDloss(parms, Q, Cap,Emp))
```


## transform to logs to get standard error for coefficients.

```{r}
logModel = glm(log(Q)~log(Cap)+log(Emp))
summary(logModel)
A = exp(logModel$coefficients[1])
alpha = logModel$coefficients[2]
beta = logModel$coefficients[3]
print(paste('A = ',A,' alpha = ', alpha,' beta =', beta))

# CD loss at these values:
parms=c(A, alpha, beta)
print(paste('CD loss: ', CDloss(parms, Q, Cap, Emp)))
```

$$log(Q) = 0.54 + 0.23*log(K) * 0.77*log(L)$$


## exponentiating
$$Q= exp(0.54)*K^{0.23}*L^{0.77}$$




## Questions:
1 what are the estimated values of A, alpha and beta
  a. A, the efficiency term is is rounded to 0.54, alpha is 0.23, and beta is 0.77. We found the values by unpacking the the antilog of the final log equation. 
  $$ A = 1.72 , \alpha= 0.23, \beta= 0.77$$
  
2. confidence intervals on A, alpha, beta?
    a. Using the information from the linear model, which produced estimates for the three variables, we can calculate CI for each variable easily. estimatedvalue plusorminus standard error

confidence interval for $\A$:log(A)$\pm$ 1.96*0.15 
  exponentiate: exp(log(A))= A$\pm$*0.15
  confidence interval for log(A): log(A)=$\pm$1.96*0.15
  log(A)=0.54
  CI for $A$ = (exp(0.54), exp(0.82)) = (1.28, 2.27)
  UCL= 2.1
  LCL= 0.46
  
confidence interval for $\alpha$:0.23$\pm$ 1.96*.03
  CI for $\alpha$ =  (exp(0.23), exp(0.82)) = (1.25 , 2.27)
  UCL= 3.52
  LCL= -1.02
  
confidence interval for $\beta$:0.77$\pm$ 1.96*.03
  CI for $\beta$ = (exp(0.77), exp(0.82)) = (2.15 , 2.27)
  UCL= 4.42
  LCL=-0.12
  
3. is production at constant return to scale? If production is increasing or decreasing return to scale, which input is most expensive
  Producion is at increasing return to scale, and employment is the most expensive input because it has the highest range of possible values according to its estimated Upper confidence level. 
  
4. What does the values of A mean??
  $A$=exp(0.54) 
  $A$=1.72
  A is a constant variable that can be studied to see if there has been a significant impact as other variables are manipulated/changed. 
